\large{\bf Looking forwards and backwards through locally regulated populations }

### 2022-04-14 forwards-and-backwards.tex
Update on Overleaf.

\begin{remark}
In Appendix \ref{sec: Markov Mapping Theorem Application},
applying Markov Mapping Theorem,
we will show that the spatial projection of $(\xi^{\theta}_t)_{t \geq 0}$ defined as 
$$(\eta^{\theta}(t))_{t \geq 0}
:=\left(\frac{1}{N}\sum\limits_{(X(t),U(t))\in \xi^{\theta}(t)} \delta_{X(t)}\right)_{t \geq 0}$$
has the same distribution (as stochastic processes)
as $(\eta^{\theta}_t)_{t \geq 0}$
in Definition \ref{Def: Rescaled Models}.
As a result, we use the same notation $\eta$
to denote these two identically distributed processes.
On the other hand,
we reserve the notation $\xi$
strictly for the spatial-level process.

Furthermore, the application of Markov Mapping Theorem implies that 
particle levels are, conditional on $\eta$,
i.i.d and uniformly distributed between $[0, N]$.

Theorem A.10 in \cite{kurtz/rodrigues:2011}
states that convergence of $(\eta^{\theta}_t)_{t \geq 0}$
to a conditionally Poisson system $(\xi_t)_{t \geq 0}$
with Cox measure $\Xi_t \times N$
is equivalent to convergence of
$(\eta^{\theta}_t)_{t \geq 0}$ to $(\Xi_t)_{t \geq 0}$.
As a result, it suffices for us to focus on
tightness of $(\eta^{\theta}_t)_{t \geq 0}$
and spatial convergence of $(\eta^{\theta}_t)_{t \geq 0}$
as $\theta \to \infty$.

@@

To prove this,
we apply Taylor's series expansion
and make use of the Martingale Problem characterisation
of $\eta^{\theta}_T$.
We have
\begin{equation}
\begin{aligned}
&\mathbb{E}\left[F_f(\eta^{\theta}_{t+s})-F_f(\eta^{\theta}_t)\bigg| \mathcal{F}_t \right]\\
=& F'(\langle f, \eta^{\theta}_t \rangle )
\mathbb{E}\left[
                \langle f, \eta^{\theta}_{t+s}\rangle
                    -\langle f, \eta^{\theta}_t\rangle 
                    \bigg| \mathcal{F}_t
         \right]
+\frac{1}{2}F''(\langle f, \eta^{\theta}_t \rangle )
\mathbb{E}\left[
                \left(
                    \langle f, \eta^{\theta}_{t+s}\rangle -\langle f, \eta^{\theta}_t\rangle
                \right)^2
                \bigg| \mathcal{F}_t
        \right]\\
&+\frac{1}{6}F'''(\langle f, \eta^{\theta}_{t+s'}\rangle)\mathbb{E}\left[
                \left(
                    \langle f, \eta^{\theta}_{t+s}\rangle -\langle f, \eta^{\theta}_t\rangle
                \right)^3
                \bigg| \mathcal{F}_t
        \right]\\
=& F'(\langle f, \eta^{\theta}_t \rangle )\\
& \times  \int_{0}^{t}\bigg\{
        \int\left(   \int \theta
     \left(
        f(z)r\big(z,\rho_{r}*\eta^{\theta}_{s}(z)\big)
            -f(x)r\big(x,\rho_{r}*\eta^{\theta}_{s}(x)\big)
            \right)
                    q_\theta(x,dz)\right)
        \\ & \qquad \qquad \qquad \qquad {} \times
                            \gamma\big(x,\rho_{\gamma}*\eta^{\theta}_{s}(x)\big)
                            \eta^{\theta}_{s}(dx)
     \\ & \qquad \qquad {} +\int 
     f(x)F\big(x,\rho_{F}*\eta^{\theta}_{s}(x)\big)
     \eta^{\theta}_{s}(dx)
    \bigg\} ds\\
=+\frac{1}{2}F''(\langle f, \eta^{\theta}_t \rangle )\\
&\times \frac{\theta}{N} \int_{0}^{t}\bigg\{
\langle \gamma\big(x,\rho_{\gamma}*\eta^{\theta}_{s}(x)\big)
\int f^2(y)r\big(y,\rho_{r}*\eta^{\theta}_{s}(y)\big)q_\theta(x,dy) 
,\eta^{\theta}_{s}(dx)\rangle\\
& \qquad \qquad +\langle \mu\big(x,\eta^{\theta}_{s}\big)f^2(x) 
,\eta^{\theta}_{s}(dx)\rangle \bigg\}ds.
\end{aligned}
\end{equation}
\end{proof}

>>>>

In Appendix \ref{sec: Markov Mapping Theorem Application},
we will show that
if we define $\Gamma (d\textbf{u})$
as the probability kernel that corresponds to
assigning each particle an i.i.d. uniform $[0,N]$-level,
and for any $\xi \in \mathcal{N}_f(\mathbb{R}^d \times [0,\infty))$, we denote
$$\hat{f}(\xi)=\int f(\xi) \Gamma (d\textbf{u})$$ 
to be the spatial test functions with averaged level,
then the process
\begin{equation}
\begin{aligned}
M^{\theta}_t:=&\hat{f}(\xi_t)-\hat{f}(\xi_0)-\int_{0}^{t}\int   A^{\theta}f(\xi_s)\Gamma(d\textbf{u})ds\\
:=&\hat{f}(\xi_t)-\hat{f}(\xi_0)-\int_{0}^{t}   \Pgen^{\theta}\hat{f}(\xi_s)ds
\end{aligned}    
\end{equation}
is a martingale.

By the Markov Mapping Theorem,
there exists a solution $(\xi^{\theta}_t)_{t \geq 0}$
to the Martingale Problem with generator $A^{\theta}$.
Furthermore,
the spatial projection of $(\xi^{\theta}_t)_{t \geq 0}$
shares the same law of $(\eta^{\theta}(t))_{t \geq 0}$,
i.e.
$$(\eta^{\theta}(t))_{t \geq 0}
\sim \left(\frac{1}{N}\sum\limits_{(X(t),U(t))\in \xi^{\theta}(t)} \delta_{X(t)}\right)_{t \geq 0},$$
and conditional on 
$(\eta^{\theta}(t))_{t \geq 0}$,
the levels of particles in $(\xi^{\theta}_t)_{t \geq 0}$
are i.i.d. uniformly distributed between $[0,N]$.

Therefore, we use the same notation $\eta$
in our spatial model
to denote the spatial projection of $\xi$,
and we reserve the notation $\xi$
strictly for the spatial-level process.

Theorem A.10 in \cite{kurtz/rodrigues:2011}
states that convergence of $(\eta^{\theta}_t)_{t \geq 0}$
to a conditionally Poisson system $(\xi_t)_{t \geq 0}$
with Cox measure $\Xi_t \times \Lambda$
is equivalent to convergence of
$(\eta^{\theta}_t)_{t \geq 0}$ to $(\Xi_t)_{t \geq 0}$.
As a result, it suffices for us to focus on
tightness of $(\eta^{\theta}_t)_{t \geq 0}$
and spatial convergence of $(\eta^{\theta}_t)_{t \geq 0}$
as $\theta \to \infty$.

@@

To prove this,
we apply Taylor's series expansion
and make use of the Martingale Problem characterisation
of $\eta^{\theta}_T$.
We have
\begin{equation}
\begin{aligned}
&\mathbb{E}\left[F_f(\eta^{\theta}_{t+s})-F_f(\eta^{\theta}_t)\bigg| \mathcal{F}_t \right]\\
=& F'(\langle f, \eta^{\theta}_t \rangle )
\mathbb{E}\left[
                \langle f, \eta^{\theta}_{t+s}\rangle
                    -\langle f, \eta^{\theta}_t\rangle 
                    \bigg| \mathcal{F}_t
         \right]
+\frac{1}{2}F''(\langle f, \eta^{\theta}_t \rangle )
\mathbb{E}\left[
                \left(
                    \langle f, \eta^{\theta}_{t+s}\rangle -\langle f, \eta^{\theta}_t\rangle
                \right)^2
                \bigg| \mathcal{F}_t
        \right]\\
&+\frac{1}{6}F'''(\langle f, \eta^{\theta}_{t+s'}\rangle)\mathbb{E}\left[
                \left(
                    \langle f, \eta^{\theta}_{t+s}\rangle -\langle f, \eta^{\theta}_t\rangle
                \right)^3
                \bigg| \mathcal{F}_t
        \right]\\
=& F'(\langle f, \eta^{\theta}_t \rangle )\\
& \times  \int_{0}^{t}\bigg\{
        \int\left(   \int \theta
     \left(
        f(z)r\big(z,\rho_{r}*\eta^{\theta}_{s}(z)\big)
            -f(x)r\big(x,\rho_{r}*\eta^{\theta}_{s}(x)\big)
            \right)
                    q_\theta(x,dz)\right)
        \\ & \qquad \qquad \qquad \qquad {} \times
                            \gamma\big(x,\rho_{\gamma}*\eta^{\theta}_{s}(x)\big)
                            \eta^{\theta}_{s}(dx)
     \\ & \qquad \qquad {} +\int 
     f(x)F\big(x,\rho_{F}*\eta^{\theta}_{s}(x)\big)
     \eta^{\theta}_{s}(dx)
    \bigg\} ds\\
&+\frac{1}{2}F''(\langle f, \eta^{\theta}_t \rangle )\\
&\quad \times \frac{\theta}{N} \int_{0}^{t}\bigg\{
\langle \gamma\big(x,\rho_{\gamma}*\eta^{\theta}_{s}(x)\big)
\int f^2(y)r\big(y,\rho_{r}*\eta^{\theta}_{s}(y)\big)q_\theta(x,dy) 
,\eta^{\theta}_{s}(dx)\rangle\\
& \qquad \qquad +\langle \mu\big(x,\eta^{\theta}_{s}\big)f^2(x) 
,\eta^{\theta}_{s}(dx)\rangle \bigg\}ds\\
&+\frac{1}{6}F^{(3)}(\langle f, \eta^{\theta}_{t+s'}\rangle)\mathbb{E}\left[
                \left(
                    \langle f, \eta^{\theta}_{t+s}\rangle -\langle f, \eta^{\theta}_t\rangle
                \right)^3
                \bigg| \mathcal{F}_t
        \right]
\end{aligned}
\end{equation}
for some $s' \in [0,s]$.
\end{proof}
